Appendix

### 2023-10-19 manuscript/review/authorsResponse.qmd
last changes before sent for internal round

author:
  - name: Olalla Díaz-Yáñez*,
    attributes:
        corresponding: true
  - name: Yannek Käber,
  - name: Tim Anders,
  - name: Friedrich Bohn,
  - name: Kristin H. Braziunas,
  - name: Josef	Brůna,
  - name: Rico Fischer,
  - name: Samuel M. Fischer,
  - name: Jessica Hetzer,
  - name: Thomas Hickler,
  - name: Christian	Hochauer,
  - name: Manfred J. Lexer,
  - name: Heike	Lischke,
  - name: Mats Mahnken,
  - name: Paola	Mairota,
  - name: Ján Merganič,
  - name: Katarina Merganičová,
  - name: Tobias Mette,
  - name: Marco Mina$,
  - name: Xavier Morin,
  - name: Werner Rammer,
  - name: Christopher P.O. Reyer,
  - name: Simon	Scheiter,
  - name: Daniel Scherrer,
  - name: Harald Bugmann

@@

# Internal instructions

-   underlined comments are comments that still need attention

# Journal formatting (ToDo list)

@@

Thank you for taking the time to review our manuscript. We appreciate your thoughtful comments and have carefully considered each of them. We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of our manuscript.

@@

[Pending]{.underline}

@@

Response to comment 20:

@@

Response to comment 21:

@@

Response to comment 24:

@@

We would like to thank Reviewer #2 for this positive overall assessment and the level of detail of the review. We have taken care to go through all suggestions made in the PDF, and have changed/improved the manuscript in most instances; however these suggestions are too numerous to be dealt with on an individual basis here.

@@

Response to comment 5:

@@

Response to comment 6:

@@

Response to comment 14:

We are not complete sure of the reviewers point here. If the reviewer is referring to the fact that the most important processes will be those included in the model, it is true that you can only asses those processes included or how including or not another processes impact regeneration. This was also the idea with considering a large pool of models, that we could assess how their structural complexity, or the inclusion/exclusion of a certain process affects the results we are seeing. This also goes very much in line with our research recommendations presented in the discussion where we proposed that improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives.

### [COMMENT 15]{.underline}

@@

[ToDo find a reference! Any suggestions?]{.underline}

### [COMMENT 16]{.underline}

@@

[ToDo find a reference!Any suggestions?]{.underline}

@@

Response to comment 19:

@@

### [COMMENT 20]{.underline}

@@

[Is there a way to explain this shortly \@ Yannek ??]{.underline}

@@

Response to comment 21:

We agree with the reviewer that permanent plots such as those present in National Forest Inventories could potentially be a good source of information to obtain regeneration data. However these often are collected from managed sites, do not capture ingrowth below 10cm, and are hard to harmonized between different inventory strategies.

@@

Response to comment 41:

@@

Response to comment 42:

@@

Response to comment 43:

We have change the terminology.

@@

Response to comment 41:

@@

We agree with the reviewer that the Shannon index is very well known index and its equation as well. However we think is still important that we define how we did calculate the Shannon index based on the available data we had as we had some internal discussion on how to do this and the meaning of this index in this context, we though the reader will benefit from a more detailed explanation. In any case the exact details of how we did calculate this index based on the simulation outputs are also available in the software supplementary materials, in the script figures.R L42-78. We also thank the reviewer for the suggestion of using the effective number of species instead of the Shannon index, it is indeed a good alternative, however we have decided to leave it with the SI.

@@

###

@@

Response to comment 52:

@@

We agree with the reviewers statement that we need the "right amount" of complexity which considers the most relevant processes. however there are limited studies looking at this aspect in models of forest dynamics and even more limited looking at this aspect in the regeneration related processes in models of forest dynamics. We still think it is important to discuss this issue in the context of our findings even though this was not the solely focus of our research.

###

@@

Response to comment 56:

@@

Response to comment 57:

In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for Global level models as they usually lack dispersal between cells.

@@

Response to comment 58:

@@

Response to comment 60:

@@

"Models included, their characteristics and simulation strategies used in this study. Where Scale refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *the* mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated , *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details. Further information can be found on the individual model reports available in the supplementary materials."

>>>>

author:
  - name: Olalla Díaz-Yáñez*,
    attributes:
        corresponding: true
  - name: Yannek Käber,
  - name: Tim Anders,
  - name: Friedrich Bohn,
  - name: Kristin H. Braziunas,
  - name: Josef	Brůna,
  - name: Rico Fischer,
  - name: Samuel M. Fischer,
  - name: Jessica Hetzer,
  - name: Thomas Hickler,
  - name: Christian	Hochauer,
  - name: Manfred J. Lexer,
  - name: Heike	Lischke,
  - name: Mats Mahnken,
  - name: Paola	Mairota,
  - name: Ján Merganič,
  - name: Katarina Merganičová,
  - name: Tobias Mette,
  - name: Marco Mina,
  - name: Xavier Morin,
  - name: Werner Rammer,
  - name: Christopher P.O. Reyer,
  - name: Simon	Scheiter,
  - name: Daniel Scherrer,
  - name: Harald Bugmann

@@

# Internal instructions \[this section will be deleted\]

-   underlined comments are comments that still need attention
-   I will delete most of the quoted text and add Line numbers, all the L\[XX-XX\] will be added in the last round after all the changes are done

## Journal formatting (ToDo list)

@@

Thank you for taking the time to review our manuscript. We have made changes to the manuscript based on your feedback and believe that these changes have improved the overall quality of our manuscript.

@@

The objective of running the different t-tests were to allow us for discussing irrefutable differences between models and observations. The t-test, as employed in our study, is not intended to serve as definitive proof of our findings. Rather, it serves as a practical tool for managing the extensive number of comparisons in our study. As such, we intend to include a statement in the manuscript:

"Note that the t-test does not present conclusive evidence for a specific hypotheses, but rather facilitates managing the extensive number of comparisons discussed in our study."

Regarding the other alternative approaches suggested by the reviewer: We can not do a paired t-test as the for paired samples it is necessary that you have the same number of subjects in both groups (observed and simulated) which is not the case. We tested the possibility of creating a GLME using as random effect the site and as a explanatory variables all the variables that we were interested in (model complexity, feedback, model type and scale). This model approach did not work as the combination as including some models having the same values in the combination of type and scale (factors did not have 2 or more levels).

-   "This more accurately answer your question about if any of these fixed explanatory variables can predict over or under representation of recruitment. " how to respond to this?

Account for Type I error rates and adjust your p-values accordingly

@@

**Response to comment 20:**

@@

**Response to comment 21:**

@@

@@

We would like to thank Reviewer 2 for this positive overall assessment and the level of detail of the review. We have taken care to go through all suggestions made in the PDF, and have changed/improved the manuscript in most instances. Our responses to each of the comments and the concrete actions we have taken are detailed below.

@@

@@

@@

We are not completely sure of the reviewers point here. If the reviewer is referring to the fact that the most important processes will be those included in the model, it is true that you can only asses those processes included or how including or not another processes impact regeneration. This was also the idea with considering a large pool of models, that we could assess how their structural complexity, or the inclusion/exclusion of a certain process affects the results we are seeing. This also goes very much in line with our research recommendations presented in the discussion where we proposed that improvement of the regeneration modules is implemented as additional features that can be traced back, as done here for the variants of ForClim and ForCEEPS, and that model complexity and structure must always be connected with modeling objectives.

@@

We have added a reference here.

@@

We have added a reference here.

@@

@@

@@

We have added information to briefly explain the general approach.

@@

**Response to comment 21:**

We agree with the reviewer that permanent plots such as those present in National Forest Inventories could potentially be a good source of information to obtain regeneration data. However these often are collected from managed sites, do not capture ingrowth below 10cm, and are hard to harmonize between different inventory strategies.

@@

Response to comment 38:

@@

Response to comment 39:

@@

Response to comment 40:

@@

**Response to comment 41:**

@@

We agree with the reviewer that the Shannon index is very well known index and its equation as well. However we think is still important that we define how we did calculate the Shannon index based on the available data we had as we had some internal discussion on how to do this and the meaning of this index in this context, we thought the reader will benefit from a more detailed explanation. In any case the exact details of how we did calculate this index based on the simulation outputs are also available in the software supplementary materials, in the script figures.R L42-78. We also thank the reviewer for the suggestion of using the effective number of species instead of the Shannon index, it is indeed a good alternative, however we have decided to leave it with the SI.

@@

@@

**Response to comment 52:**

@@

We agree with the reviewers statement that we need the "right amount" of complexity which considers the most relevant processes. However there are limited studies looking at this aspect in models of forest dynamics and even more limited looking at this aspect in the regeneration related processes in models of forest dynamics. We still think it is important to discuss this issue in the context of our findings even though this was not the solely focus of our research.

@@

**Response to comment 56:**

@@

**Response to comment 57:**

In this sentence we are discussing how the sampling strategy presented in the study protocol might have represented a disadvantage for landscape level models due to the limited spatial scale, but how this might not be the same for global level models as they usually lack dispersal between cells.

@@

**Response to comment 58:**

@@

**Response to comment 60:**

@@

"Models included, their characteristics and simulation strategies used in this study. Where *Scale* refers to the models scale, *Type* to the models type, *Population structure* to the model population approach, *Feedback* to the inclusion or not of adult tree feedback to the level and species composition of regeneration, *Approach* to the model regeneration module approach, *Start from* to the stage from which trees are recruited, *Complexity,* to *t*he mean regeneration formulation as estimated in (Bugmann & Seidl, 2022), *Species* to which species from the original eleven species were simulated, *Runtime (spin up) \[years\]* to the number of years used per model in their simulations for the spin up period and regeneration sampling; *Climate data type (use of years)* to the type of data used in the simulations and how it was used, and Reference provides the link to the publication with more model details. Further information can be found on the individual model reports available in the supplementary materials."

### 2023-10-19 manuscript/supplementaryRW.Rmd
last changes before sent for internal round

TreeMig simulates directly the mean dynamics of entire stands, but based
on the variability of light conditions within the stand [distribution
based approach, Lischke, 1998, aggregation. To come up with the required
200 random patches of 1 ha, temporal and spatial variability intrinsic
to TreeMig were combined. Temporal variability was given by sampling
from each decade of the last 100 years of the simulation, i.e. years
900-1000. For each of theses 10 decades, 20 grid cells were sampled from
the simulated grid, reflecting spatial variability, results from
demographic stochasticity and short-range spatial interactions

>>>>

TreeMig simulates directly the mean dynamics of entire stands, but based
on the variability of light conditions within the stand (distribution based approach, cf. Lischke et al. 1998). To come up with the required
200 random patches of 1 ha, temporal and spatial variability intrinsic
to TreeMig were combined. Temporal variability was given by sampling
from each decade of the last 100 years of the simulation, i.e. years
900-1000. For each of theses 10 decades, 20 grid cells were sampled from
the simulated grid, reflecting spatial variability, results from
demographic stochasticity and short-range spatial interactions
